---
title: "Psychy Facility Data"
output: html_document
date: '2022-05-18'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/phsta/Documents/Psych_Hosp")
library(dplyr)
library(tidyverse)
library(ggplot2)
library(usmap)
```

## Introduction

## Data Description

## Loading Data

```{r readData Main}
load("Data/2019 data/nmhss_2019_puf_r.rdata")
psych_data <- PUF
rm(PUF)
```

## Looking at Washington State


```{r waState}
psych_data %>%
  filter(LST == "WA")
```

## Writing the variable names to a csv file to manipulate seperately in excel for a bit

```{r write names}
var_labels <- lapply(psych_data, attr, "label")
psych_names <- t(data.frame(var_labels))

write.csv(t(data.frame(var_labels)), file = "psych_data_vars.csv")
```

```{r read vars with mods}
psych_names <- read.csv("psych_data_vars_modified.csv")
names(psych_names) <- c("psy_df_name", "label", "Type.1", "Type.2") 
```

## Playing around with state displays

```{r state displays}
num_facilities <- psych_data %>% group_by(LST) %>%
  rename(state = LST) %>%
  summarize(num_fac = n())

plot_usmap(data = num_facilities, values = "num_fac") +
  scale_fill_continuous(name = "Number of Facilities (2019)", label = scales::comma) +
  labs(title = "Number of Facilities by State (2019)") +
  theme(legend.position = "right")
  
```

Now let's say we want to get per capita information state by state. We are going to need an additional dataset, merge it with num_facilities tbl, create a per capita variable, then plot that. 

```{r percapita map}
state_pops <- read.csv("Data/demo data/State_pops.csv", header = TRUE)

num_facilities <- merge(num_facilities, state_pops) %>%
  mutate(per_10000_r = (num_fac/population)*10000)

plot_usmap(data = num_facilities, values = "per_10000_r") +
  scale_fill_continuous(name = "Number of Facilities per 10,000 (2019)", label = scales::comma) +
  labs(title = "Number of Facilities by State by 10,000 people (2019)") +
  theme(legend.position = "right")
```

Next, I want to try and capture the number of mental health service providers in a state per person with mental illness and serious mental illness. To do this, I will need to interact with the NSDUH, a giant file. I will try and pull the full file in and then select a smaller number of columns order to reduce the burden on R. I will want to capture mental illness (AMIYR_U), serious mental illness (SMIYR_U), serious to moderate mental illness (SMMIYR_U), and weights. Note that the mental illness values are all imputed (see codebook for additional information). These also don't line up perfectly for the percentages on the front page of the website for 2019, but are only 0.1% off. I wonder if it's the difference of the actual data vs the PUF?


```{r NSDUH attempt}
load("Data/NSDUH_2020.Rdata")

NSDH_mi <- NSDUH %>% select
```

Ok, actually the state information for the NSDUH is suppressed. I think there is a lot of interesting information in the NSDUH to share with Shelby for potential work, but it isn't super releavnt at the moment. Instead, I am going to pull the percentages of mental illess from the online dataset, store it in a csv, load that in and build the tables from that. Even that's a bit tricky though. Rates of mental illness per state are only available for individuals 18+ on the website. I chose to pull information from the census's imputation for 2019. I believe this was conducted before the 2020 census results were available, so it may not be entirely accurate. Would probably be worth checking whether there are major changes between imputed 2019 and observed 2020. 
